---
title: "Gape Tracker calibration code"
author: "Luke Miller"
output: 
  word_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 5
---

```{r knitSetup, echo=FALSE}
showcode = FALSE # Set true to print out all R code in final document

# Comment out one of the two localpath lines below depending on what computer you're on
localpath = '../../data/'  # Luke's path
#localpath = '../data/'  # Lauren's path

```

# Calibration walk-through

## Fall 2022
We calibrated individual Hall effect gape sensors attached to BivalveBit
boards. The Allegro A1395 Hall effect sensor was used
for all of these sensors. 

We calibrated using a set of neodymium magnets (Digikey part
number 469-1005-ND), so that all sensors could use the same magnet type. 
Magnets were covered in adhesive-lined heat shrink to waterproof them, and
the polarity of each magnet was marked with a silver paint pen to facilitate
proper installation on the oyster. 

Calibration was done using a KD Scientific Model 200 syringe pump to move
the magnet known distances away from the Hall effect sensor. The syringe
pump was set to move the magnet in 0.3175 mm steps, starting near the 
Hall sensor and moving away from it, up to a maximum of 12.3825 mm from the
starting point, which we arbitrarily call 0.0 mm. A calibration program,
`Hall_calibration_routine.ino` was uploaded to the BivalveBit
boards for calibration. The program wrote csv files to the onboard micro SD
card. Each file contained the following columns:

* `Serial` - BivalveBit board serial number (i.e. SN102, SN103)
* `Trial` - the trial number for this calibration series. Usually 1-10,
  unless trials had to be deleted due to syringe pump issues.
* `Distance.mm` - the assumed distance the syringe pump carriage had
  moved from the initial starting position (defined as 0.0mm, although 
  there was always a gap between the magnet and the Hall effect sensor).
* `Reading` - the analog-to-digital converter value from the
  BivalveBit's ATmega4808 onboard A-D (10 bit maximum precision, 0-1023)
  The magnets and sensors were oriented so that readings were low when 
  the magnet was closest to the sensor, and increased towards 512 as
  the magnet moved away from the sensor. In practice the minimum
  values were around 10-15 before the sensor saturated, and the
  maximum was around 450-480 when the magnet had moved at least 12.3
  mm away from the sensor.
  
Calibration files were arranged in folders by date, and each file name
followed the pattern `SNxxx_YYYYMMDD_HHMM_CALIB.csv`, with xxx = the board's
serial number and `YYYYMMDD_HHMM` marking the date and time of the trial.

Each calibration trial was set with the Hall sensor attached to the 
syringe pump so that the bottom of the sensor board was pointing
towards the magnet (the same as it would be oriented on the shell of the
animal). The magnet was attached to the moving carriage portion of
the syringe pump. The magnet was moved near the sensor to establish
a starting value (usually between 40 and 400 counts). The software
then began by taking a set of 4 Hall effect readings and averaging
them, storing that average in the csv file, and then moving the
carriage 0.3175 mm before taking the next reading. This cycle would
repeat until the maximum distance of 12.3 mm was reached, after
which the carriage was reset near the Hall sensor for the next
trial. The starting orientation of the magnet and Hall sensor would
be altered slightly for each trial, so that a sample of different
alignments could be measured. The goal was to cover the range of
possible alignments that might occur when the sensor and magnet were
glued to the shell of the animal. 

On occasion the moving carriage of the syringe pump would not fully engage with
the lead screw, and the correct movement would not be achieved until the
carriage snapped into place. This would be noted by the operator, and the data
for that trial would be later removed from the data file.  

Once the calibration trials were completed, the data were saved to
the project folder. In
cases where one or more trials in a file were suspect due to the
carriage not engaging properly, the rows for that trial were
deleted from the data file, so the resulting csv files should only
contain known-good trials. 




## Using the calibration data



```{r gapeFunctions, echo=showcode}

# A set of functions originally taken from MussleTracker_data_proc4.R to 
# handle the gape data
		
		################################################################################
# exciseTempHall: A function to excise questionable rows of data from the 1-Hz 
# temperature and Hall effect sensor data frame
# Inputs:
#	temphall: a data frame of temperature and hall effect data from the
#			Mussel Tracker raw data files
#	excisedf: a data frame of start and end times to ignore in the data set
# Output:
# 	temphall: a modified version of the input data frame with the questionable
#			rows replaced with NAs. 
		
#' A function to excise questionable rows of data from the 1-Hz 
#' temperature and Hall effect sensor data frame
#' @param temphall A data frame of temperature and hall effect data from the
#'			Mussel Tracker raw data files
#' @param excisedf A data frame of start and end times to ignore in the data set 
#' @return temphall A data frame with 2 columns, Start and End, which contain 
#' the row indices for the first good and last good values in each contiguous 
#' run of numeric values. 
		
		exciseTempHall = function(temphall, excisedf) {
	# Find the rows in excisedf that have the same serial number as the current
	# board, and have time points to remove for each channel and sensor combo
	# Start with Ch1 Temperature
	matches = which(excisedf$SerialNumber == levels(temphall$SerialNumber)[1] &
					excisedf$Mussel == 'Ch1' & excisedf$Sensor == 'Temp')
	if (length(matches) > 0){
		for (i in 1:length(matches)){
			startrow = which.min(abs(temphall$DateTime - 
									excisedf$StartIgnore[matches[i]]))
			endrow = which.min(abs(temphall$DateTime - 
									excisedf$EndIgnore[matches[i]]))
			# blank out the suspect rows of data
			temphall$Temp1[startrow:endrow] = NA
		}
	}
	# Do the same for Channel 2 temperature
	matches = which(excisedf$SerialNumber == levels(temphall$SerialNumber)[1] &
					excisedf$Mussel == 'Ch2' & excisedf$Sensor == 'Temp')
	if (length(matches) > 0){
		for (i in 1:length(matches)){
			startrow = which.min(abs(temphall$DateTime - 
									excisedf$StartIgnore[matches[i]]))
			endrow = which.min(abs(temphall$DateTime - 
									excisedf$EndIgnore[matches[i]]))
			# blank out the suspect rows of data
			temphall$Temp2[startrow:endrow] = NA
		}
	}
	# Next remove bad data from Hall sensor Ch1
	matches = which(excisedf$SerialNumber ==levels(temphall$SerialNumber)[1] &
					excisedf$Mussel == 'Ch1' & excisedf$Sensor == 'Hall')
	if (length(matches) > 0){
		for (i in 1:length(matches)){
			startrow = which.min(abs(temphall$DateTime - 
									excisedf$StartIgnore[matches[i]]))
			endrow = which.min(abs(temphall$DateTime - 
									excisedf$EndIgnore[matches[i]]))
			# blank out the suspect rows of data
			temphall$Hall1[startrow:endrow] = NA
		}
	}
	# And remove bad data from Hall sensor Ch2
	matches = which(excisedf$SerialNumber == levels(temphall$SerialNumber)[1] &
					excisedf$Mussel == 'Ch2' & excisedf$Sensor == 'Hall')
	if (length(matches) > 0){
		for (i in 1:length(matches)){
			startrow = which.min(abs(temphall$DateTime - 
									excisedf$StartIgnore[matches[i]]))
			endrow = which.min(abs(temphall$DateTime - 
									excisedf$EndIgnore[matches[i]]))
			# blank out the suspect rows of data
			temphall$Hall2[startrow:endrow] = NA
		}
	}
	temphall # return data frame as output
} # End of exciseTempHall function

###############################################################################	
# A function to determine where there are gaps of missing data, so that we can 
# apply a filter to the good chunks of data. Most filters fail if there are NAs
# present, and our remaining gaps in the data are big enough that they probably
# shouldn't be smoothed-over anyhow. 
#' A function to find the start and end indices of numeric data in a vector. 
#' Useful for dealing with timeseries that have gaps, and operations that need
#' to operate the separate chunks of contiguous numeric values
#' @param value A vector of numeric values, with NAs 
#' @return A data frame with 2 columns, Start and End, which contain the row
#' indices for the first good and last good values in each contiguous run of
#' numeric values. 	
gapBounds <- function (values){
	# This function returns a data frame of 2 columns, start + end, that give
	# the row indices of the start and end of each run of good data in the 
	# input vector. It should handle data streams that start with NAs and
	# should handle streams that end with NAs. 
	
	# Use run length encoding function to find NA gaps
	gaps = rle(is.na(values))
	# If the first entry in gaps is TRUE, this indicates that the data started
	# with NA values.
	if (gaps$values[1] == TRUE) {
		startNA = TRUE	# Data started with NAs
	} else {	
		startNA = FALSE	 # Data started with real values
	}
	# If the last entry in gaps is TRUE, the data end on NAs
	if (gaps$values[length(gaps$values)] == TRUE){
		endNA = TRUE	# Data ends with NAs
	} else {
		endNA = FALSE	# Data ends with real values
	}
	
	# The number of gaps with value == TRUE is the number of good runs of data
	# and gaps with value == FALSE are the starts of runs of NAs. This will
	# get the number of FALSE (negated to TRUE) values in gaps. A dataset
	# ending in good data will have a last entry in gaps of FALSE. 
	numgaps = sum(!gaps$values)
	
	# Create the output data frame
	results = data.frame(Start = integer(numgaps), End = integer(numgaps))
	
	if (!startNA) {
		# If startNA == FALSE (started on good data, the first entry should 
		# always be 1 (1st row) 
		# If there are no gaps, the
		# contents of gaps$lengths will just be the length of the values vector
		results$Start[1] = 1
		results$End[1] = gaps$lengths[1]
	} else if (startNA){
		# However, if the dataset starts with NAs, the first entry in gaps will
		# be the index of the first good data, while the 2nd entry will be the 
		# start of the next stretch of NAs
		results$Start[1] = gaps$lengths[1]+1
		results$End[1] = sum(gaps$lengths[1:2])
	}
	
	# If there is more than 1 entry in gaps$lengths, process the other gaps
	j = 2; # counter
	if (numgaps > 1){
		if (!startNA & endNA){
			# If the dataset ends on NAs (TRUE last), truncate gapind by 1
			gapind = seq(2,length(gaps$lengths)-1, by = 2)
		} else if (!startNA & !endNA) {
			# If the dataset ends on good values (FALSE last)
			gapind = seq(2,length(gaps$lengths), by = 2)
		} else if (startNA & endNA) {
			# If dataset starts on NAs (TRUE 1st) and ends on NAs (TRUE last)
			gapind = seq(3, length(gaps$lengths)-1, by = 2)
		} else if (startNA & !endNA) {
			# If dataset starts on NAs (TRUE 1st) and ends on good data 
			# (FALSE last)
			gapind = seq(3, length(gaps$lengths), by = 2)
		}	
		# Step through the rest of the gaps object to find the start and end
		# points of each stretch of good data. 
		for (i in gapind){
			nextstart = sum(gaps$lengths[1:i]) + 1
			nextend = sum(gaps$lengths[1:(i+1)])
			results$Start[j] = nextstart
			results$End[j] = nextend
			j = j + 1
		}
	} # end of if (numgaps > 1)
	results	# return the results dataframe, 2 columns Start and End
}

################################################################################
### hallFilter function
#' A function to apply a Butterworth 1st order low-pass filter to a vector
#'  of data. If there are NA values in the original data, they will be 
#'  preserved in the output vector. The filter is set to filter at 1/10 the
#' original sampling rate (which is 1Hz normally), producing an approximate
#' 10-second window of smoothing. Raw input data are initially centered on 0
#' to avoid artifacts at the start of the filter, and the output is re-centered
#' at the original starting value after the filtering is applied. 
#' @param hallData A vector of hall effect data.
#' @return A vector of filtered hall effect sensor data. 
hallFilter = function(hallData){	
	# Define a butterworth low-pass filter, 1st order, to filter at 1/10 the 
	# sampling rate (which was 1Hz)
	myfilter = butter(1,0.1,type='low', plane = 'z')
	# Find the start and end of any gaps in the data set using the gapBounds
	# function defined earlier
	mygaps = gapBounds(hallData)
	# Apply the filter to the good runs of data
	for (i in 1:nrow(mygaps)){
		# Extract the run of good data
		dats = hallData[mygaps$Start[i]:mygaps$End[i]]
		# Find the offset of the data from zero
		offset = dats[1]
		# Subtract the offset off of all of the data in this run
		# (The butterworth filter returns large transient values at the start
		# of the run if the value is much different from zero)
		dats = dats - offset
		# Call the filter routine to apply the filter
		yfiltered = signal:::filter(myfilter,dats)
		# Add the offset back on so the data are back on their original scale
		yfiltered = yfiltered + offset
		# Write the filtered data back into the data vector
		hallData[mygaps$Start[i]:mygaps$End[i]] = yfiltered
	}
	# Round the filtered data back to the nearest whole number, since these
	# represent ADC count data.
	hallData = round(hallData)
	hallData	# return vector of filtered data
}

################################################################################
# Define a non-linear function that allows you to plug in a Hall effect sensor
# count value and generate an approximate distance in mm. Supply the 
# coefficients a, b, c, based on a nls curve fit for counts as a function of
# distance. 
# In the above equation, 'a' is the effective asymptotic value
# 'b' is the range of values between the minimum observed count and the 
# asymptotic count (so a min count of 400 and asymptote of 500 gives b = 100)
# 'c' is effectively a curvature parameter. Larger values of 'c' yield a 
# smaller change in Distance as Counts increases (when back-calculating 
# Distance), particularly for low Counts values. 
#' A function to back-calculate an asymptotic function to generate a distance
#' value based on a hall effect sensor count value input.
#' 
#' @param Counts A vector of hall effect sensor values, usually between 0 and 
#' 512 (from a 10-bit analog-to-digital converter). 
#' @param a A parameter derived from an asymptotic curve fit, representing the
#' asymptotic count value
#' @param b A parameter derived from an asymptotic curve fit, representing the 
#' range from the minimum count up to the asymptotic value. 
#' @param c A parameter derived from an asymptotic curve fit, representing the
#' curvature of the fit. 
#' @returns A vector of distance values of the same length as Counts. Any input 
#' values equal or larger than the asymptotic parameter a will return Inf.

DistFromCounts <- function(Counts, a,b,c){
	dist = (log((Counts - a)/(-1*b))) / (-1*c)
	dist
}


################################################################################
### percentileRange function
# Define a function to calculate the average reading for the upper and lower
# x percentile of the hall effect data. 
#' Calculate the average of the specified lower percentile and upper percentile
#' and return the values that correspond to those means (rounded to integers).
#' 
#' @param hallData A vector of analog-to-digital convertor counts from a 
#' Hall effect sensor. The values are assumed to start low (<512) when the 
#' magnetic signal is strong (close to sensor) and increase towards 512 when
#' the magnetic signal is weak (far from sensor).
#' @param percentileLim A numeric value between 0 and 1, usually close to 0.01
#' used to define the lower and upper percentile limits (mirrored). A value of 
#' 0.01 would cause the 1st and 99th percentile sensor values to be used as the
#' lower and upper limits
#' @return A two element vector for the lower and upper values in hallData that
#' represent the lower and upper percentiles

percentileRange <- function (Hallvec, percentileLim = 0.01){
	# Remove any NA's
	temp = Hallvec[!is.na(Hallvec)]
	# Reorder the values smallest to largest
	temp = temp[order(temp)]
	# Get the index of the entry closest to percentileLim
	indx = round(percentileLim * length(temp))
	# Calculate the mean value for the lower 1% of closed valve values and
	# round down to the next integer value
	closedVal = floor(mean(temp[1:indx]))
	
	# Now do the same for the other end of the range of hall effect values
	# These would normally represent "fully open" readings near 512 if the 
	# magnet and sensor are situated so that a nearby magnet drives the signal
	# below 512. 
	indx = round((1 - percentileLim) * length(temp))
	# Calculate the mean value for the upper 1% of open valve values and round
	# up to the next integer value
	openVal = ceiling(mean(temp[indx:length(temp)]))
	result = c(closedVal, openVal)
	result # return the two values, always smallest then largest
}

################################################################################
### calcPercentGapeNLME function
# Use this function to calculate the percentage gape (0-100%)
# This calculates the baseline fully-closed and fully-open values based
# on the upper and lower 1% of hall effect sensor data for the given range
# of rows in the input dataset. 
# Supply a vector of Hall effect sensor data. This function will find any 
# gaps in the vector (NA's) and calculate a new baseline for each section of
# good data. 
#' Convert hall effect count data (integer values from analog-digital converter)
#' into percent gape values between 0 and 100\%. 
#' 
#' @param hallData A vector of analog-to-digital convertor counts from a 
#' Hall effect sensor. The values are assumed to start low (<512) when the 
#' magnetic signal is strong (close to sensor) and increase towards 512 when
#' the magnetic signal is weak (far from sensor).
#' @param cRegress A linear model object (lm) for the relationship of 'c' versus
#' the log-transform of 'b'. 
#' @param percentileLim A numeric value between 0 and 1, usually close to 0.01,
#' used to define the lower and upper percentile limits (the upper limit is 
#' mirrored from the lower limit). A value of 
#' 0.01 would cause the 1st and 99th percentile sensor values to be used as the
#' lower and upper limits.
#' @return A vector of the same length as hallData containing percentage values 
#' 0 to 100, representing the shell valve gape 
#' opening. A value of 0 is a fully closed shell (strong magnetic signal) and
#' 100 is a fully open shell (weak magnetic signal).

calcPercentGapeNLME <- function (hallData, cRegress, percentileLim = 0.01){
	require(nlme)
	# Get the row indices for the major gaps that exist now
	mygaps = gapBounds(hallData)
	# Create an empty output vector of the same length as the input data. 
	outputData = vector(mode = 'numeric', length = length(hallData))
	outputData[] = NA # Convert all the values to NA to begin with
	
	# Now for each section of the input data, calculate the percent gape. This
	# involves using the entries in mygaps to subset the input data and 
	# calculate individual percentile values for each contiguous section of
	# data. If there is a significant gap in the data (usually > 30 sec), the
	# percentages will be re-calculated for each contiguous section of data. 
	# This should accommodate any sensor/magnet re-glueing issues. 
	for (i in 1:nrow(mygaps)){
		st = mygaps$Start[i]
		end = mygaps$End[i]
		# First calculate the upper and lower 1% values of the Hall readings
		# These will be count values representing the fully closed (lowest) 
		# and fully open (highest, closest to 512) values.
		myrange = percentileRange(hallData[st:end], 
				percentileLim = percentileLim)
		
		# Now truncate count values that are outside myrange to the limits of
		# myrange. 
		rowIndices = which(hallData[st:end] < myrange[1])
		hallData[(st+rowIndices)] = myrange[1]
		rowIndices = which(hallData[st:end] > myrange[2])
		hallData[(st+rowIndices)] = myrange[2]
		
#		hallData[which(hallData[st:end] < myrange[1])] = myrange[1]
#		hallData[which(hallData[st:end] > myrange[2])] = myrange[2]

		# The count values are now constrained within a range that should 
		# encompass fully closed (0mm) to fully open (~5mm gape). Now calculate
		# the approximate gape opening based on the count values and a curve
		# fit. 
		
		# Define the three asymptotic curve parameters a,b,c using the data for
		# the current chunk of data.
		aVal = max(hallData[st:end],na.rm=TRUE)+1 # add 1 count to avoid Inf output
		bVal = max(hallData[st:end],na.rm=TRUE) - 
				min(hallData[st:end],na.rm=TRUE)
		# Calculating c requires the linear model object supplied as cRegress,
		# which will use the log-transformed 'b' value to estimate the 'c'
		# parameter
		cVal = predict(cRegress, newdata = list(b = bVal))
		
		# Use DistFromCounts function to estimate valve gape distance using the 
		# input hall effect count data at each time point and the 
		# parameters a,b,c
		temp2 = DistFromCounts( hallData[st:end], 
				a = aVal,
				b = bVal,
				c = cVal)
		
		# Replace any fully-open Inf values with the largest good max distance
		temp2[is.infinite(temp2)] = max(temp2[is.finite(temp2)])
		# Replace any negative distance values with the 0 distance value
		temp2[which(temp2 < 0)] = 0
		# Convert to percentage by dividing all distances by the maximum value
		outputData[st:end] = (temp2 / max(temp2,na.rm=TRUE)) * 100
		
		# Round the values off to a reasonable precision
		outputData[st:end] = round(outputData[st:end],1)
		
	}	
	outputData	# return the output data vector
}	# end of calcPercentGapeNLS function


################################################################################
################################################################################
# Function cRegressEstimate
# Currently this represents the best data showing how the relationship between
# magnet signal and distance is repeatable and consistent among different
# magnets (all from the same batch/part number). As such, a non-linear curve
# can be fit to the various calibration runs to generate a set of 
# coefficients a,b,c for an asymptotic curve. Then the values from a field 
# mussel can be analyzed to determine the asymptotic maximum count value
# (widest opening) and minimum count value (fully closed), which give us the
# parameters for coefficients a and b in the curve fit. Because the curvature
# coefficient 'c' of the curves is very linear with respect to the coefficient
# 'b' (range of readings), we can plug in the max vs. min range for a particular
# mussel and use that to predict the appropriate curve coefficient 'c' to use
# for that mussel. 

#' A function to generate a set of hall effect calibration data and estimate
#' the relationship between parameters 'b' and 'c', for later use in the 
#' function calcPercentGapeNLME(). 
#' @param SN A text string naming the board serial number, in the form 'SN21'
#' @param Channel numeric value naming the Hall effect sensor channel to be 
#' calibrated (expected values 0-15). 
#' @param gapeCalibFile A text path to the file that holds Hall effect
#' calibration data
#' @return cRegress A linear model object representing the best fit regression
#'  of parameter 'c' on the log-transformed parameter 'b'. 

cRegressEstimate <- function(SN, Channel, gapeCalibFile){
	require(nlme)
	# Create an identifier for the board/channel combination 'SN21Ch01'
	if (Channel < 10){
		newChannel = paste0('0',as.character(Channel))
	} else {
		newChannel = as.character(Channel)
	}
	currentID = paste0(SN,'Ch',newChannel)
	 
	# Get list of calibration data
	gapeCalibs = read.csv(gapeCalibFile)

	gapeCalibs$Serial = factor(gapeCalibs$Serial)
	gapeCalibs$Trial = factor(gapeCalibs$Trial)
	# Create a unique identifier for each board/channel combination,
	# ensuring that the Channel number is always 2 digits
	gapeCalibs$ID = factor(paste0(gapeCalibs$Serial,'Ch',
					ifelse(gapeCalibs$HallChannel<10,
							paste0('0',gapeCalibs$HallChannel),
							gapeCalibs$HallChannel)))

# For any instance where magnet was set up so that values climbed above 512
# when close to the sensor, reverse those values so they are all less than 
# 512. 
	for (i in 1:length(levels(gapeCalibs$ID))){
		if (max(gapeCalibs$Reading[gapeCalibs$ID == 
								levels(gapeCalibs$ID)[i]]) > 515){
			gapeCalibs$Reading[gapeCalibs$ID == 
							levels(gapeCalibs$ID)[i]] = 1023 - 
					gapeCalibs$Reading[gapeCalibs$ID == 
									levels(gapeCalibs$ID)[i]]
		}
	}	
	# Begin generating the asymptotic curve fit parameters for the magnetic
# hall effect sensor data. 
	gC2 = groupedData(Reading~Distance.mm|ID/Trial,data=gapeCalibs)
	

# Fit the asymptotic model and generate a set of coefficient values
# This will adjust each set of a,b,c coefs based on a random effect of
# magID and Trial. 
	
	# Fit separate curves to each group (trial) in the grouped data frame
	mod3 = nlsList(Reading~a-(b*exp(-c*Distance.mm)), 
			data = subset(gC2,subset = ID == currentID), 
			start = c(a = 460, b = 60, c = 0.3),
			control = list(minFactor = 1e-9, msMaxIter = 200, maxiter = 200))
# Extract the a,b,c coefficients and put them in a data frame
	abcvals = data.frame(a = coef(mod3)[,1], b = coef(mod3)[,2],
			c = coef(mod3)[,3])
	
# The fit between 'b' and 'c' can be linearized with a log transform of 'b'
	cRegress = lm(c~log(b), data = abcvals)  

#	plot(c~`log(b)`, data = cRegress$model) # Plot the original data
#	abline(cRegress) # fitted regression line
	
	# Return cRegress as the output
	cRegress
}  # end of cRegressEstimate function.
################################################################################

```

Once the curve parameters $a,b,$ and $c$ have been established for a particular
sensor and magnet setup, the distance of opening in millimeters (relative to the
initial closed position) can be estimated using the equation:

$Distance = \frac{log(\frac{Counts - a}{-1 \times b})}{-1 \times c}$

The original curve fits, to estimate Reading values from Distance, are 
based on the equation:

$Counts = a-(b \times \exp^{(-c \times Distance)})$

The accuracy of the estimated distance unavoidably degrades as the magnet and
Hall sensor get further apart (i.e. wider shell openings), but for the 
distances related to the initial shell opening, we can get reasonably good
(sub-millimeter) estimates of the change in distance between the magnet and
sensor (and thus between the 2 valves of the shell)

Below is a figure showing the observed and fitted estimates of the readings
for each Distance, from multiple starting points. We see good correspondence
between the fitted values and observed values, indicating that the estimated
coefficients

```{r trialCurveFitting, echo=FALSE, eval = FALSE, dpi=300,fig.width=5,fig.cap='Observed readings (open circles) and fitted values (closed circles) derived from nls curve fitting routine.'}
## Example of code to fit curves to each Trial for a given sensor, followed by
# showing the predicted values from the curve fits afterwards. 

library(nlme)
# Begin generating the asymptotic curve fit parameters for the magnetic
# hall effect sensor data. 
gC2 = groupedData(Reading~Distance.mm|Trial,data=SN20Hall02, order.groups = FALSE)
#plot(Reading~Distance.mm,data=gC2)


# Fit separate curves to each group (trial) in the grouped data frame
mod3 = nlsList(Reading~a-(b*exp(-c*Distance.mm)), 
		data = gC2, start = c(a = 460, b = 30, c = 0.3),
		control = list(minFactor = 1e-9, msMaxIter = 200))
# Extract the a,b,c coefficients and put them in a data frame
abcvals = data.frame(a = coef(mod3)[,1], b = coef(mod3)[,2],
		c = coef(mod3)[,3])

# The fit between 'b' and 'c' can be linearized with a log transform of 'b'
cRegress = lm(c~log(b), data = abcvals)  
#plot(c~log(b),data=abcvals)

# Run predict function for a set of 50 fixed distances
reps = 100
newdist = seq(0,12.3,length = reps)
res = predict(mod3, newdata = list(Distance.mm = newdist))
# Create a data frame from the results, with the Distance.mm values added
# back in and a Trial number assigned for each group of 50 distances
predlines = data.frame(Reading = res, Distance.mm= rep(newdist,10),
		Trial = factor(rep(1:10,each=reps)))
# Basic plot
plot(Reading~Distance.mm, data = predlines, type = 'n',
		main = 'SN20Hall02')
for (i in 1:length(levels(SN20Hall02$Trial))){
	# Plot observed points
	points(Reading~Distance.mm, 
			data = SN20Hall02[SN20Hall02$Trial == levels(SN20Hall02$Trial)[i],],
			pch = 21, col = i)
	# Plot observed lines
#	lines(Reading~Distance.mm,
#			data = SN20Hall02[SN20Hall02$Trial == levels(SN20Hall02$Trial)[i],],
#			pch = 21, col = i, lwd = 1.5)
	# Plot fitted points based on curve fitting routine
	points(Reading~Distance.mm, 
			data = predlines[predlines$Trial == levels(predlines$Trial)[i],],
			pch = 20, col = i)
}
legend('bottomright',legend=c('Observed','Fitted'),pch=c(21,20))

```

## Calibration results for each sensor

For these calibration trials, we set the starting distance between the 
magnet and Hall effect sensor to several different values, to account for 
the range of possible starting values we might get when the magnet and sensor
are glued to the oyster shell. The important thing here is that the shape
of the curves is consistent, changing only in steepness depending on the
initial starting distance.


```{r defineGapeCalibFile, echo=showcode, cache=FALSE}
# Give the path to the concatenated calibration file
calibDir ='../data/Calibration/201905_all_calibrations/'
gapeCalibFile = paste0(calibDir,'201905_calibrations.csv')
```

```{r plotCalibFunction, echo=showcode }

#' A function to open calibration data and plot data for a given board and channel
#' 
#' @param SN A text string naming the board serial number, in the form 'SN21'
#' @param Channel numeric value naming the Hall effect sensor channel to be 
#' calibrated (expected values 0-15). 
#' @param gapeCalibFile A text path to the file that holds Hall effect
#' calibration data
#' @param plot Logical value indicating whether a plot should be produced
#' @param plotpoints Logical value indicating whether points should be plotted
#' @param plotlines Logical value indicating whether lines should be plotted
#' @param plottext Logical value indicating whether text markers indicating the
#' Trial number should be plotted
#' @param insetplot Logical value indicating whether an inset plot showing the
#' relationship between the estimated curve parameters log(b) and c should be
#' included
#'  

plotCalib <- function(SN,Channel,gapeCalibFile, plot = TRUE, plotpoints = TRUE, 
		plotlines = TRUE, plottext = FALSE, insetplot = FALSE) {
	# Create an identifier for the board/channel combination, i.e. 'SN21Ch01'
	if (Channel < 10){
		newChannel = paste0('0',as.character(Channel))
	} else {
		newChannel = as.character(Channel)
	}
	# Create the ID for the current board & channel
	currentID = paste0(SN,'Ch',newChannel)
	# Open the concatenated file of calibration data
	gapeCalibs = read.csv(gapeCalibFile) 
	gapeCalibs$Serial = factor(gapeCalibs$Serial)
	gapeCalibs$Trial = factor(gapeCalibs$Trial)
	# Create a unique identifier for each board/channel combination, 
	# ensuring that the Channel number is always 2 digits 
	gapeCalibs$ID = factor(paste0(gapeCalibs$Serial,'Ch',
					ifelse(gapeCalibs$HallChannel<10,
							paste0('0',gapeCalibs$HallChannel),
							gapeCalibs$HallChannel)))
	# Generate the curve parameters for this board & channel calibration						
	cRegress = cRegressEstimate(SN, Channel, gapeCalibFile)	
	# Subset out the chosen board & channel's data
	dat = gapeCalibs[gapeCalibs$ID == currentID,]
	dat = droplevels(dat)
	
	if(plot){
		par(fig = c(0,1,0,1)) # Set up default main plot extent
		ylims = range(dat$Reading) 
		xlims = range(dat$Distance.mm) 
		plot(Reading~Distance.mm,data = dat, type = 'n',
				ylim = ylims,
				las = 1,
				ylab = 'Raw AD Counts',
				xlab = 'Distance, mm',
				main = currentID)
		rect(par()$usr[1],par()$usr[3],par()$usr[2],par()$usr[4],col='grey70')
		grid()
		box()
		
		# Plot 
		for (i in 1:length(levels(dat$Trial))){
			if (plotpoints){
				points(Reading~Distance.mm, 
						data = dat[dat$Trial == levels(dat$Trial)[i],],
						pch = 21, col = i)				
			}
			if (plotlines){
				lines(Reading~Distance.mm,
						data = dat[dat$Trial == levels(dat$Trial)[i],],
						pch = 21, col = i, lwd = 1.5)
			}
			if (plottext){
				text(Reading~Distance.mm, 
						data = dat[dat$Trial == levels(dat$Trial)[i],],
					labels = dat[dat$Trial == levels(dat$Trial)[i],'Trial'],
						col = 'black')
			}
		}
		if (insetplot) {
			par(fig = c(0.4,0.95,0.15,0.75), new = TRUE)
			regressR2 = summary(cRegress)$adj.r.squared
			# Plot the original data
			plot(c~`log(b)`, data = cRegress$model, las = 1, pch = 20)
			rect(par()$usr[1],par()$usr[3],par()$usr[2],par()$usr[4],col='grey80')
			points(c~`log(b)`, data = cRegress$model, las = 1, pch = 20)
			text(c~`log(b)`, data = cRegress$model, 
					labels = names(cRegress$fitted.values), adj = -0.2)
			abline(cRegress) # fitted regression line
			mylabel = bquote(italic(R)^2 == .(format(regressR2, digits = 3)))
			legend('topleft',legend=mylabel, bty = 'n', cex = 0.8)
		}
	}
}

```
		

### SN21 Calibrations

```{r SN21Hall00,echo=showcode, cache=FALSE,fig.cap='SN21Hall00',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN21',Channel = 0,gapeCalibFile, plottext = FALSE, 
		insetplot = TRUE) 
```

```{r SN21Hall01,echo=showcode, cache=FALSE,fig.cap='SN21Hall01',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN21',Channel = 1,gapeCalibFile, plottext = FALSE, 
		insetplot = TRUE)
```

```{r SN21Hall02,echo=showcode,fig.cap='SN21Hall02',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN21',Channel = 2,gapeCalibFile, plottext = FALSE,  
		insetplot = TRUE)
```

```{r SN21Hall03,echo=showcode, cache=FALSE,fig.cap='SN21Hall03',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN21',Channel = 3,gapeCalibFile, plottext = FALSE, 
		insetplot = TRUE)
```

```{r SN21Hall04,echo=showcode, cache=FALSE,fig.cap='SN21Hall04',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN21',Channel = 4,gapeCalibFile, plottext = FALSE, 
		insetplot = TRUE)
```

```{r SN21Hall08,echo=showcode,fig.cap='SN21Hall08',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN21',Channel = 8,gapeCalibFile, plottext = FALSE, 
		insetplot = TRUE)
```

```{r SN21Hall09,echo=showcode, cache=FALSE,fig.cap='SN21Hall09',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN21',Channel = 9,gapeCalibFile, plottext = FALSE, 
		insetplot = TRUE)
```

```{r SN21Hall10,echo=showcode, cache=FALSE,fig.cap='SN21Hall10',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN21',Channel = 10,gapeCalibFile, plottext = FALSE, 
		insetplot = TRUE)
```

```{r SN21Hall11,echo=showcode, cache=FALSE,fig.cap='SN21Hall11',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN21',Channel = 11,gapeCalibFile, plottext = FALSE,  
		insetplot = TRUE)
```

```{r SN21Hall12,echo=showcode, cache=FALSE,fig.cap='SN21Hall12',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN21',Channel = 12,gapeCalibFile, plottext = FALSE, 
		insetplot = TRUE)
```


### SN20 Calibrations


```{r SN20Hall00,echo=showcode, cache=FALSE,fig.cap='SN20Hall00',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN20',Channel = 0,gapeCalibFile, plottext = FALSE, 
		insetplot = TRUE)
```

```{r SN20Hall01,echo=showcode, cache=FALSE,fig.cap='SN20Hall01',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN20',Channel = 1,gapeCalibFile, plottext = FALSE,
		insetplot = TRUE)
```

```{r SN20Hall02,echo=showcode, cache=FALSE,fig.cap='SN20Hall02',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN20',Channel = 2,gapeCalibFile, plottext = FALSE,
		insetplot = TRUE)
```

```{r SN20Hall03,echo=showcode, cache=FALSE,fig.cap='SN20Hall03',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN20',Channel = 3,gapeCalibFile, plottext = FALSE,
		insetplot = TRUE)
```

```{r SN20Hall04,echo=showcode, cache=FALSE,fig.cap='SN20Hall04',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN20',Channel = 4,gapeCalibFile, plottext = FALSE,
		insetplot = TRUE)
```


```{r SN20Hall08,echo=showcode, cache=FALSE,fig.cap='SN20Hall08',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN20',Channel = 8,gapeCalibFile, plottext = FALSE, 
		insetplot = TRUE)
```

```{r SN20Hall09,echo=showcode, cache=FALSE,fig.cap='SN20Hall09',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN20',Channel = 9,gapeCalibFile, plottext = FALSE, 
		insetplot = TRUE)
```


```{r SN20Hall10,echo=showcode, cache=FALSE,fig.cap='SN20Hall10',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN20',Channel = 10,gapeCalibFile, plottext = FALSE,
		insetplot = TRUE)
```


```{r SN20Hall11,echo=showcode, cache=FALSE,fig.cap='SN20Hall11',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN20',Channel = 11,gapeCalibFile, plottext = FALSE, 
		insetplot = TRUE)
```


```{r SN20Hall12,echo=showcode, cache=FALSE,fig.cap='SN20Hall12',fig.width=5,fig.height=5,dpi=600}
plotCalib(SN = 'SN20',Channel = 12,gapeCalibFile, plottext = FALSE, 
		insetplot = TRUE)
```



```{r testingLoad, eval=FALSE, echo=FALSE}
# Use this chunk to examine a single calibration file by providing the
# directory name and file name that you want to look at. 
# Load some data
fdir = "../data/Calibration/20190514_calibration/"
fname = 'Hall01_CALIB_00_SN20.csv'
# Assemble the directory and file name
gapeCalibFile = paste0(fdir,fname)
# Get the Serial number from the file name, match pattern 'SNxx'
matchVal = regexpr(pattern='SN[[:digit:]]{2}', text = fname)
# Extract the magnet ID as text
SN = substr(x=fname,
		start = matchVal, 
		stop = matchVal + (attr(matchVal,'match.length')-1))
# Get channel number, match pattern 'Hallxx'
matchVal = regexpr(pattern='Hall[[:digit:]]{2}', text = fname)
Channel = as.numeric(substr(x=fname,
		start = matchVal+4,
		stop = matchVal + (attr(matchVal,'match.length')-1)))
# Produce the calibration plot and c~log(b) plot
plotCalib(SN,Channel,paste0(fdir,fname),plottext=TRUE,insetplot=TRUE)
```